---
title: 'Figure 2: Core analysis'
author: "Sofie Zacho Vestergaard"
date: "2024-01-03"
output: html_document
---
```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r}
library(ampvis2)
library(tidyverse)
library(data.table)
library(patchwork)
library(viridis)
```

# load data
```{r}
d <- amp_load(otutable = 'data/ASV_table_combined.tsv',
              taxonomy = 'data/ASVs_combined_MiDAS5.2.R1.sintax',
              metadata = 'data/metadata.xlsx',
              pruneSingletons = T) # remove singletons

# merge replicates
d_rep <- amp_merge_replicates(d, 
                              merge_var = "SampleID_replicates", 
                              round = 'up')

d_rep$metadata <- d_rep$metadata %>%
  rename(SampleID = SampleID_replicates)

d_rep_filt_as <- amp_subset_samples(d_rep,
                                    SampleContent %in% 'AS',
                                    minreads = 10000, removeAbsentOTUs = F)

set.seed(22) # set seed to make rarefaction reproducible
# rarefy
d_rare_as <- amp_rarefy(d_rep_filt_as, rarefy = 10000) # 10 000

#normalise
dasn <- normaliseTo100(d_rare_as)
```


##1)ASV

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Create taxonomy data frame (for later)
tax <- dasn$tax[1:7]
tax$Kingdom <- gsub("k__","",tax$Kingdom)
tax$Phylum <- gsub("p__","",tax$Phylum)
tax$Class <- gsub("c__","",tax$Class)
tax$Order <- gsub("o__","",tax$Order)
tax$Family <- gsub("f__","",tax$Family)
tax$Genus <- gsub("g__","",tax$Genus)
tax$Species <- gsub("s__","",tax$Species)
tax$OTU <- rownames(tax)
tax <- distinct(tax)

dlong <- amp_export_long(
    dasn,
    metadata_vars = "SampleSite", #data are aggregated at plant level 
    tax_levels = c("OTU"))
```

###CORE
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#group up and summarise for core taxa
  core <- dlong[, .(sum = sum(count)), by = c("OTU", "SampleSite")]   
  data.table::setorderv(core, c("SampleSite", "sum"), order = -1)
  #calculate proportion % abundance of each ASV
  core[, Speciesprop := sum / sum(sum) * 100, by = "SampleSite"]
  #calculate how many times given ASV is observed in plants (>0)
  core <- core[OTU!=""] #remove rows where no species (unclassified ASVs)
  core[, nObs := sum(sum > 0), by = OTU]
  #calculate how many times (in how many plants) ASV is >0.1% abundant
  core[, nCore := sum(Speciesprop > 0.1), by = OTU]
  #add column with sum of all the reads
  core[, totalReads := sum(sum)]
    core_summary <- core[
    , 
    .(nSpecies = uniqueN(OTU)),
    by = .(nObs, nCore)]

```


```{r}
###Retrive core data
#subset ASVs present in the core
core_loose <- core[(core$nCore >= n_distinct(core$SampleSite)*0.2),] %>% #loose core (20% plants)
	group_by(OTU) %>%
	summarise(mean_abu = mean(Speciesprop)) %>%
           arrange(desc(mean_abu))
core_loose <- core_loose[ core_loose$OTU != "", ]
core_loose$OTU <- gsub("s__","",core_loose$OTU)
core_loose[,"V1V3"] <- 2
core_loose[,"Category"] <- "loose core"

core_general <- core[(core$nCore >= n_distinct(core$SampleSite)*0.5),] %>% #general core (50% plants)
	group_by(OTU) %>%
	summarise(mean_abu = mean(Speciesprop)) %>%
           arrange(desc(mean_abu))
core_general <- core_general[ core_general$OTU != "", ]
core_general$OTU <- gsub("s__","",core_general$OTU)
core_general[,"V1V3"] <- 3
core_general[,"Category"] <- "general core"

core_strict <- core[(core$nCore >= n_distinct(core$SampleSite)*0.8),] %>% #strict core (80% plants)
	group_by(OTU) %>%
	summarise(mean_abu = mean(Speciesprop)) %>%
           arrange(desc(mean_abu))
core_strict <- core_strict[ core_strict$OTU != "", ]
core_strict$OTU <- gsub("s__","",core_strict$OTU)
core_strict[,"V1V3"] <- 4
core_strict[,"Category"] <- "strict core"
```

###CRAT



```{r}
### Plot for conditional abundant genera
  
#group up and summarise for conditional abundant taxa (CAT)
crat <- dlong[, .(sum = sum(count)), by = c("OTU", "SampleSite")]   
  setorderv(core, c("SampleSite", "sum"), order = -1)
  #calculate proportion % abundance of each ASV
  crat[, Speciesprop := sum / sum(sum) * 100, by = "SampleSite"]
  #calculate how many times given ASV is observed in plants (>0)
  crat <- crat[OTU!=""]


 crat <- crat[!OTU %in% c(core_loose$OTU, core_general$OTU, core_strict$OTU)] # Remove core ASVs 
 
  crat[, nObs := sum(sum > 0), by = OTU]
  #calculate how many times (in how many plants) ASV is >1% abundant
  crat[, nCA := sum(Speciesprop > 1), by = OTU]
  #add column with sum of all the reads
  crat[, totalReads := sum(sum)]
    crat_summary <- crat[,.(nSpecies = uniqueN(OTU)), by = .(nObs, nCA)]
    
```




```{r}
###Retrive CRAT data
#subset ASVs present in the core
core_crat <- crat[(crat$nCA >= 1)] %>%
	group_by(OTU) %>%
	summarise(mean_abu = mean(Speciesprop)) %>%
           arrange(desc(mean_abu))
core_crat <- core_crat[ core_crat$OTU != "", ]
core_crat[,"V1V3"] <- 1
core_crat[,"Category"] <- "CRAT"
```

###combine core&crat data
```{r}
#Merge core data for v13
core_loose <- filter(core_loose, !(OTU %in% core_general$OTU)) #loose core (>20%) should not include general core (>50%)
core_general <- filter(core_general, !(OTU %in% core_strict$OTU))#general core (>50%) should not include strict core (>80%)

core2 <- rbind(core_strict[, c(1,4)], core_general[, c(1,4)]) %>%
  rbind(., core_loose[, c(1,4)]) %>%
  rbind(., core_crat[, c(1,4)]) %>%
  merge(., tax[,c("Genus", "Species", "OTU")], by="OTU", all.x=TRUE)
```

###Stats
```{r}
#nr of ASVs in each categories
core2 %>% 
  group_by(Category) %>% 
  distinct(OTU, .keep_all = T) %>% 
  count()

#check if by accident some ASVs have been countend twice. It should return empty
core2 %>% 
  filter(duplicated(core2$OTU) == T) %>% 
  arrange(OTU, Category) %>% 
  View()

```

###export list core-crat
```{r}
openxlsx::write.xlsx(core2,'output/core&crat_ASV.xlsx', rowNames = F)
```

##2)Genus
```{r, echo=FALSE, message=FALSE, warning=FALSE}

dlong_gen <- amp_export_long(
    dasn,
    metadata_vars = "SampleSite", 
    tax_levels = c("Genus"))
```

###CORE
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#group up and summarise for core taxa
  core_gen <- dlong_gen[, .(sum = sum(count)), by = c("Genus", "SampleSite")]   
  data.table::setorderv(core_gen, c("SampleSite", "sum"), order = -1)
  #calculate proportion % abundance of each ASV
  core_gen[, GenusProp := sum / sum(sum) * 100, by = "SampleSite"]
  #calculate how many times given ASV is observed in plants (>0)
  core_gen <- core_gen[Genus!=""] #remove rows where no species (unclassified ASVs)
  core_gen[, nObs := sum(sum > 0), by = Genus]
  #calculate how many times (in how many plants) ASV is >0.1% abundant
  core_gen[, nCore := sum(GenusProp > 0.1), by = Genus]
  #add column with sum of all the reads
  core_gen[, totalReads := sum(sum)]
    core_gen_summary <- core_gen[
    , 
    .(nGenus = uniqueN(Genus)),
    by = .(nObs, nCore)]

```



```{r}
###Retrive core data
#subset ASVs present in the core
core_gen_loose <- core_gen[(core_gen$nCore >= n_distinct(core_gen$SampleSite)*0.2),] %>% #loose core (20% plants)
	group_by(Genus) %>%
	summarise(mean_abu = mean(GenusProp)) %>%
           arrange(desc(mean_abu))
core_gen_loose <- core_gen_loose[ core_gen_loose$Genus != "", ]
core_gen_loose$Genus <- gsub("g__","",core_gen_loose$Genus)
core_gen_loose[,"V1V3"] <- 2
core_gen_loose[,"Category"] <- "loose core"

core_gen_general <- core_gen[(core_gen$nCore >= n_distinct(core_gen$SampleSite)*0.5),] %>% #general core (50% plants)
	group_by(Genus) %>%
	summarise(mean_abu = mean(GenusProp)) %>%
           arrange(desc(mean_abu))
core_gen_general <- core_gen_general[ core_gen_general$Genus != "", ]
core_gen_general$Genus <- gsub("g__","",core_gen_general$Genus)
core_gen_general[,"V1V3"] <- 3
core_gen_general[,"Category"] <- "general core"

core_gen_strict <- core_gen[(core_gen$nCore >= n_distinct(core_gen$SampleSite)*0.8),] %>% #strict core (80% plants)
	group_by(Genus) %>%
	summarise(mean_abu = mean(GenusProp)) %>%
           arrange(desc(mean_abu))
core_gen_strict <- core_gen_strict[ core_gen_strict$Genus != "", ]
core_gen_strict$Genus <- gsub("g__","",core_gen_strict$Genus)
core_gen_strict[,"V1V3"] <- 4
core_gen_strict[,"Category"] <- "strict core"
```


###CRAT

```{r}
#group up and summarise for conditional abundant taxa (CAT)
crat_gen <- dlong_gen[, .(sum = sum(count)), by = c("Genus", "SampleSite")]   
  setorderv(core_gen, c("SampleSite", "sum"), order = -1)
  #calculate proportion % abundance of each ASV
  crat_gen[, GenusProp := sum / sum(sum) * 100, by = "SampleSite"]
  #calculate how many times given ASV is observed in plants (>0)
  crat_gen <- crat_gen[Genus!=""]

  crat_gen$Genus <- gsub("g__","",crat_gen$Genus)#remove genus string so that the following filtering can work
  
 crat_gen <- crat_gen[!Genus %in% c(core_gen_loose$Genus, core_gen_general$Genus, core_gen_strict$Genus)] # Remove core ASVs 
 
  
  crat_gen[, nObs := sum(sum > 0), by = Genus]
  #calculate how many times (in how many plants) ASV is >1% abundant
  crat_gen[, nCA := sum(GenusProp > 1), by = Genus]
  #add column with sum of all the reads
  crat_gen[, totalReads := sum(sum)]
  crat_gen_summary <- crat_gen[,.(nGenus = uniqueN(Genus)), by = .(nObs, nCA)]
```



```{r}
###Retrive CRAT data
#subset Genera present in the core
core_crat_gen <- crat_gen[(crat_gen$nCA >= 1)] %>%
	group_by(Genus) %>%
	summarise(mean_abu = mean(GenusProp)) %>%
           arrange(desc(mean_abu))
core_crat_gen <- core_crat_gen[ core_crat_gen$Genus != "", ]
core_crat_gen[,"V1V3"] <- 1
core_crat_gen[,"Category"] <- "CRAT"
```

###combine core&crat data
```{r}
#Merge core data for v13
core_gen_loose <- filter(core_gen_loose, !(Genus %in% core_gen_general$Genus)) #loose core (>20%) should not include general core (>50%)
core_gen_general <- filter(core_gen_general, !(Genus %in% core_gen_strict$Genus))#general core (>50%) should not include strict core (>80%)

core_gen2 <- rbind(core_gen_strict, core_gen_general) %>%
  rbind(., core_gen_loose) %>%
  rbind(., core_crat_gen) 

```

###Stats
```{r}
#check nr of genera in each category
core_gen2 %>% 
  group_by(Category) %>% 
  distinct(Genus, .keep_all = T) %>% 
  count()
```

###export list core-crat
```{r}
openxlsx::write.xlsx(core_gen2,'output/core&crat_genus.xlsx', rowNames = F)
```

##3)Species

```{r, echo=FALSE, message=FALSE, warning=FALSE}

dlong_sp <- amp_export_long(
    dasn,
    metadata_vars = "SampleSite", #data are aggregated at plant level (abundance recalculated at plant level)
    tax_levels = c("Species"))
```

###CORE
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#group up and summarise for core taxa
  core_sp <- dlong_sp[, .(sum = sum(count)), by = c("Species", "SampleSite")]   
  data.table::setorderv(core_sp, c("SampleSite", "sum"), order = -1)
  #calculate proportion % abundance of each ASV
  core_sp[, SpeciesProp := sum / sum(sum) * 100, by = "SampleSite"]
  #calculate how many times given ASV is observed in plants (>0)
  core_sp <- core_sp[Species!=""] #remove rows where no species (unclassified ASVs)
  core_sp[, nObs := sum(sum > 0), by = Species]
  #calculate how many times (in how many plants) ASV is >0.1% abundant
  core_sp[, nCore := sum(SpeciesProp > 0.1), by = Species]
  #add column with sum of all the reads
  core_sp[, totalReads := sum(sum)]
    core_sp_summary <- core_sp[
    , 
    .(nSpecies = uniqueN(Species)),
    by = .(nObs, nCore)]

```



```{r}
###Retrive core data
#subset ASVs present in the core
core_sp_loose <- core_sp[(core_sp$nCore >= n_distinct(core_sp$SampleSite)*0.2),] %>% #loose core (20% plants)
	group_by(Species) %>%
	summarise(mean_abu = mean(SpeciesProp)) %>%
           arrange(desc(mean_abu))
core_sp_loose <- core_sp_loose[ core_sp_loose$Species != "", ]
core_sp_loose$Species <- gsub("s__","",core_sp_loose$Species)
core_sp_loose[,"V1V3"] <- 2
core_sp_loose[,"Category"] <- "loose core"

core_sp_general <- core_sp[(core_sp$nCore >= n_distinct(core_sp$SampleSite)*0.5),] %>% #general core (50% plants)
	group_by(Species) %>%
	summarise(mean_abu = mean(SpeciesProp)) %>%
           arrange(desc(mean_abu))
core_sp_general <- core_sp_general[ core_sp_general$Species != "", ]
core_sp_general$Species <- gsub("s__","",core_sp_general$Species)
core_sp_general[,"V1V3"] <- 3
core_sp_general[,"Category"] <- "general core"

core_sp_strict <- core_sp[(core_sp$nCore >= n_distinct(core_sp$SampleSite)*0.8),] %>% #strict core (80% plants)
	group_by(Species) %>%
	summarise(mean_abu = mean(SpeciesProp)) %>%
           arrange(desc(mean_abu))
core_sp_strict <- core_sp_strict[ core_sp_strict$Species != "", ]
core_sp_strict$Species <- gsub("s__","",core_sp_strict$Species)
core_sp_strict[,"V1V3"] <- 4
core_sp_strict[,"Category"] <- "strict core"
```

###CRAT

```{r}
### Plot for conditional abundant genera
  
#group up and summarise for conditional abundant taxa (CAT)
crat_sp <- dlong_sp[, .(sum = sum(count)), by = c("Species", "SampleSite")]   
  setorderv(core_sp, c("SampleSite", "sum"), order = -1)
  #calculate proportion % abundance of each ASV
  crat_sp[, SpeciesProp := sum / sum(sum) * 100, by = "SampleSite"]
  #calculate how many times given ASV is observed in plants (>0)
  crat_sp <- crat_sp[Species!=""]

  crat_sp$Species <- gsub("s__","",crat_sp$Species)#remove genus string so that the following filtering can work
  
 crat_sp <- crat_sp[!Species %in% c(core_sp_loose$Species, core_sp_general$Species, core_sp_strict$Species)] # Remove core ASVs 

  crat_sp[, nObs := sum(sum > 0), by = Species]
  #calculate how many times (in how many plants) ASV is >1% abundant
  crat_sp[, nCA := sum(SpeciesProp > 1), by = Species]
  #add column with sum of all the reads
  crat_sp[, totalReads := sum(sum)]
  crat_sp_summary <- crat_sp[,.(nSpecies = uniqueN(Species)), by = .(nObs, nCA)]
```



```{r}
###Retrive CRAT data
#subset Genera present in the core
core_crat_sp <- crat_sp[(crat_sp$nCA >= 1)] %>%
	group_by(Species) %>%
	summarise(mean_abu = mean(SpeciesProp)) %>%
           arrange(desc(mean_abu))
core_crat_sp <- core_crat_sp[ core_crat_sp$Species != "", ]
core_crat_sp[,"V1V3"] <- 1
core_crat_sp[,"Category"] <- "CRAT"

```

###combine core&crat data
```{r}
#Merge core data for v13
core_sp_loose <- filter(core_sp_loose, !(Species %in% core_sp_general$Species)) #loose core (>20%) should not include general core (>50%)
core_sp_general <- filter(core_sp_general, !(Species %in% core_sp_strict$Species))#general core (>50%) should not include strict core (>80%)

core_sp2 <- rbind(core_sp_strict, core_sp_general) %>%
  rbind(., core_sp_loose) %>%
  rbind(., core_crat_sp) %>%
  merge(., tax[,c("Genus","Species")], by="Species", all.x=TRUE) %>% 
  unique()

```

###Stats

```{r}
#nr of ASVs in each categories
core_sp2 %>% 
  group_by(Category) %>% 
  distinct(Species, .keep_all = T) %>% 
  count()
```

###export list core-crat
```{r}
openxlsx::write.xlsx(core_sp2,'output/core&crat_species.xlsx', rowNames = F)
```

#Summary core&CRAT

##import core&crat data
```{r}
corecrat_asv <- readxl::read_excel("output/core&crat_ASV.xlsx") %>% 
  rename(ASV = OTU)
corecrat_gen <- readxl::read_excel("output/core&crat_genus.xlsx")
corecrat_sp <- readxl::read_excel("output/core&crat_species.xlsx")
```


AS long format at ASV level per plant
```{r}

#hetamp: extract the abundance of all ASVs in all plants on average
as_asv_txt <- amp_heatmap(dasn,
            tax_show = 100000000000, 
            group_by = c("SampleSite"),
            tax_aggregate = "OTU",
            tax_add = c("Genus", "Species"),
            normalise = F,
            textmap = T
            )

#saveRDS(as_asv_txt, file = 'output/AS_txt_heat_ASV.rds')


```

merge long format with core-genus/species/ASV dataframes
```{r}
as_asv_txt2 <-as_asv_txt %>%
  mutate(tax = rownames(as_asv_txt)) %>%
  separate(tax, sep="; ", into=c("Genus","Species","ASV")) %>%
  mutate(Genus = ifelse(grepl("ASV|unmatched",Genus),"Unclassified",Genus)) %>%
  mutate(Species = ifelse(grepl("ASV|unmatched",Species),"Unclassified",Species)) %>%
  
  merge(corecrat_gen[,c("Genus", "Category")], by="Genus", all.x = TRUE) %>%
  mutate(Genus_type = ifelse(Genus=="Unclassified","Unclassified",Category)) %>%
  mutate(Genus_type = ifelse(is.na(Genus_type),"Other",Genus_type)) %>%
  select(-Category) %>% 
  
  merge(corecrat_sp[,c("Species", "Category")], by="Species", all.x = TRUE) %>%
  mutate(Species_type = ifelse(Species=="Unclassified","Unclassified", Category)) %>%
  mutate(Species_type = ifelse(is.na(Species_type),"Other",Species_type)) %>%
  select(-Category) %>% 
  
  merge(corecrat_asv[,c("ASV", "Category")], by="ASV", all.x = TRUE) %>% 
  rename(ASV_type = Category) %>% 
  mutate(ASV_type = ifelse(is.na(ASV_type), "Other", ASV_type)) %>% 
  mutate(ASV_type = case_when(ASV_type == "loose core" ~ "Loose core",
                                ASV_type == "general core" ~ "General core",
                                ASV_type == "strict core" ~ "Strict core",
                                TRUE ~ ASV_type)) %>% 
  mutate(Species_type = case_when(Species_type == "loose core" ~ "Loose core",
                                Species_type == "general core" ~ "General core",
                                Species_type == "strict core" ~ "Strict core",
                                TRUE ~ Species_type)) %>% 
  mutate(Genus_type = case_when(Genus_type == "loose core" ~ "Loose core",
                                Genus_type == "general core" ~ "General core",
                                Genus_type == "strict core" ~ "Strict core",
                                TRUE ~ Genus_type))

# put on numbers of each group (e.g. how many spp in general core)
as_asv_txt_number_ASV <- as_asv_txt2 %>%  
  count(ASV_type) %>% 
  rename(Number_ASVs = n) 
as_asv_txt_number_SPP <- as_asv_txt2 %>%  group_by(Species) %>% 
  count(Species_type) %>%
  ungroup() %>% count(Species_type) %>% 
  rename(Number_spp = n) %>% 
  filter(!Species_type %in% 'Unclassified')
as_asv_txt_number_genus <- as_asv_txt2 %>%  group_by(Genus) %>% 
  count(Genus_type) %>%
  ungroup() %>% count(Genus_type) %>% 
  rename(Number_gen = n) %>% 
  filter(!Genus_type %in% 'Unclassified')



# make long format
as_asv_txt2 <- as_asv_txt2 %>%
  gather("Aalbæk":"Vojens", key="Plant", value="RRA")

View(head(as_asv_txt2, 10))


```

##ASV
```{r}
#summary per plant
summary_asv <- as_asv_txt2 %>% 
  group_by(Plant,ASV_type) %>%
  summarise(cum_rra_asv = sum(RRA)) %>%
  mutate(ASV_type = factor(ASV_type, levels=c("Other", "CRAT", "Loose core", "General core", "Strict core")))

p_asv_byplant <- ggplot(data = summary_asv, aes(x = Plant, y=cum_rra_asv, fill=ASV_type)) +
geom_bar(stat = "identity") +
  labs(y = "Cumulative read abundance [%]",
        fill = "Core group",
        tag = 'A') +
    scale_fill_manual(values = rev(c("#D7191C", "#FDAE61", "#FFFFBF", "#ABD9E9", "grey")))+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, colour = "black", hjust = 1, vjust = 0.5, size = 11),
        axis.title.x = element_blank(), axis.title.y = element_text(size = 13),
        legend.position = "bottom",
        ) + scale_y_continuous(expand = c(0,0)) 

#summary across plant
summary_asv_avg <- summary_asv %>% 
  group_by(ASV_type) %>%
  summarise(mean_cum_rra_asv = mean(cum_rra_asv)) %>%
  mutate(ASV_type = factor(ASV_type, levels=c("Other", "CRAT", "Loose core", "General core", "Strict core"))) %>% 
  mutate(Avg = "All WWTPs") %>% 
  left_join(., as_asv_txt_number_ASV, by = 'ASV_type')  %>%
  mutate(ASV_type = factor(ASV_type, levels=c("Other", "CRAT", "Loose core", "General core", "Strict core")))

p_asv_acrossplant <- ggplot(data = summary_asv_avg, aes(x = Avg, y=mean_cum_rra_asv, fill=ASV_type)) +
geom_bar(stat = "identity") +
  geom_text(data = summary_asv_avg, aes(label = paste0(round(mean_cum_rra_asv, digits = 1), ' %\n(', Number_ASVs,')')),
             size = 3, position = position_stack(vjust = 0.5), lineheight = 0.72) +
  labs( fill = "Core group") +
    scale_fill_manual(
      values = rev(c("#D7191C", "#FDAE61", "#FFFFBF", "#ABD9E9", "grey")))+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, colour = "black", hjust = 1, vjust = 0.5, size = 11),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
        ) + scale_y_continuous(expand = c(0,0)) 

#combine species plot
p_summ_asv <- (p_asv_byplant + p_asv_acrossplant + theme(axis.title.y = element_blank()) ) +
  scale_x_discrete(expand = c(0,0)) +
  plot_layout(ncol = 2, nrow = 1,
    widths = c(2, 0.1),
    guides = "collect") & 
  theme(legend.position = "none") #+
  
p_summ_asv2 <- p_summ_asv + plot_annotation(
) & theme(
          plot.tag = element_text(face = 'bold', size = 18))
```

##Species
```{r}
#summary per plant
summary_sp <- as_asv_txt2 %>% 
  group_by(Plant,Species_type) %>%
  summarise(cum_rra_sp = sum(RRA)) %>%
  mutate(Species_type = factor(Species_type, levels=c("Unclassified", "Other", "CRAT", "Loose core", "General core", "Strict core")))

p_sp_byplant <- ggplot(data = summary_sp, aes(x = Plant, y=cum_rra_sp, fill=Species_type)) +
geom_bar(stat = "identity") +
  labs(y = "Cumulative read abundance [%]",
        fill = "Core group",
        tag = 'B') +
    scale_fill_manual(values = rev(c("#D7191C", "#FDAE61", "#FFFFBF", "#ABD9E9", "grey", "#f3f6f4")))+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, colour = "black", hjust = 1, vjust = 0.5, size = 11),
        legend.position = "bottom",
        axis.title.x = element_blank(), axis.title.y = element_text(size = 13)
        ) + scale_y_continuous(expand = c(0,0)) 

#summary across plant
summary_sp_avg <- summary_sp %>% 
  group_by(Species_type) %>%
  summarise(mean_cum_rra_sp = mean(cum_rra_sp)) %>%
  mutate(Species_type = factor(Species_type, levels=c("Unclassified", "Other", "CRAT", "Loose core", "General core", "Strict core"))) %>% 
  mutate(Avg = "All WWTPs") %>% 
  left_join(., as_asv_txt_number_SPP, by = 'Species_type')  %>%
  mutate(Species_type = factor(Species_type, levels=c("Unclassified","Other", "CRAT", "Loose core", "General core", "Strict core")))

p_sp_acrossplant <- ggplot(data = summary_sp_avg, aes(x = Avg, y=mean_cum_rra_sp, fill=Species_type)) +
geom_bar(stat = "identity") +
  geom_text(data = summary_sp_avg, aes(label = case_when(Species_type == 'Unclassified' ~ paste0(round(mean_cum_rra_sp, digits = 1), ' %'),TRUE ~ paste0(round(mean_cum_rra_sp, digits = 1), ' %\n(', Number_spp,')'))),
             size = 3, position = position_stack(vjust = 0.5), lineheight = 0.72) +
  labs( fill = "Core group") +
    scale_fill_manual(#palette="RdYlBu", na.value = "grey"
      values = rev(c("#D7191C", "#FDAE61", "#FFFFBF", "#ABD9E9", "grey", "#f3f6f4")))+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, colour = "black", hjust = 1, vjust = 0.5, size = 11),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
        ) + scale_y_continuous(expand = c(0,0)) 
#combine species plot
p_summ_sp <- (p_sp_byplant + p_sp_acrossplant + theme(axis.title.y = element_blank()) ) +
  scale_x_discrete(expand = c(0,0)) +
  plot_layout(ncol = 2, nrow = 1,
    widths = c(2, 0.1),
    guides = "collect") & 
  theme(legend.position = "none") 
  
p_summ_sp2 <- p_summ_sp + plot_annotation(
) & theme(plot.title = element_text(face = 'bold'),
          plot.tag = element_text(face = 'bold', size = 18))

```


##Genus
```{r}
#summary per plant
summary_gen <- as_asv_txt2 %>% 
  group_by(Plant,Genus_type) %>%
  summarise(cum_rra_gen = sum(RRA)) %>%
  mutate(Genus_type = factor(Genus_type, levels=c("Unclassified", "Other", "CRAT", "Loose core", "General core", "Strict core"))) 

p_gen_byplant <- ggplot(data = summary_gen, aes(x = Plant, y=cum_rra_gen, fill=Genus_type)) +
geom_bar(stat = "identity") +
  labs(y = "Cumulative read abundance [%]",
        fill = "Core group",
        tag = 'C') +
    scale_fill_manual(values = rev(c("#D7191C", "#FDAE61", "#FFFFBF", "#ABD9E9", "grey", "#f3f6f4")))+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, colour = "black", hjust = 1, vjust = 0.5, size = 11),
        legend.position = "bottom",
        axis.title.x = element_blank(), axis.title.y = element_text(size = 13)
        ) + guides(fill = guide_legend(override.aes = list(color = "black", size = 0.3))) + scale_y_continuous(expand = c(0,0)) 

#summary across plant
summary_gen_avg <- summary_gen %>% 
  group_by(Genus_type) %>%
  summarise(mean_cum_rra_gen = mean(cum_rra_gen)) %>%
  mutate(Avg = "All WWTPs")%>% 
  left_join(., as_asv_txt_number_genus, by = 'Genus_type')  %>%
  mutate(Genus_type = factor(Genus_type, levels=c("Unclassified","Other", "CRAT", "Loose core", "General core", "Strict core")))

p_gen_acrossplant <- ggplot(data = summary_gen_avg, aes(x = Avg, y=mean_cum_rra_gen, fill=Genus_type)) +
geom_bar(stat = "identity") +
  geom_text(data = summary_gen_avg, aes(label = case_when(Genus_type == 'Unclassified' ~ paste0(round(mean_cum_rra_gen, digits = 1), ' %'), TRUE ~ paste0(round(mean_cum_rra_gen, digits = 1), ' %\n(', Number_gen,')'))),
             size = 3, position = position_stack(vjust = 0.5), lineheight = 0.72) +
  labs( fill = 'Core group') +
    scale_fill_manual(#palette="RdYlBu", na.value = "grey"
      values = rev(c("#D7191C", "#FDAE61", "#FFFFBF", "#ABD9E9", "grey", "#f3f6f4")))+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, colour = "black", hjust = 1, vjust = 0.5, size = 11),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
        ) + guides(fill = guide_legend(override.aes = list(color = "black", size = 0.3)))+ scale_y_continuous(
                      expand = c(0,0)) 
#combine species plot
p_summ_gen <- (p_gen_byplant + p_gen_acrossplant + theme(axis.title.y = element_blank()) ) +
  scale_x_discrete(expand = c(0,0))+
  plot_layout(ncol = 2, nrow = 1,
    widths = c(2, 0.1),
    guides = "collect") & 
  theme(legend.position = "bottom", legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)) 
  
p_summ_gen2 <- p_summ_gen + plot_annotation(
) & theme(plot.title = element_text(face = 'bold'),
          plot.tag = element_text(face = 'bold',size = 18)) 

```
##combo plot
```{r}
library(ggpubr)
p <- ggarrange(p_summ_asv2,
          p_summ_sp2,
          p_summ_gen2,
          ncol = 1,
          nrow = 3,
          heights = c(1, 1, 1.2))
ggsave("/Users/kl42gg/Downloads/Fig02_Core&CRAT_wNUM.png",
        p,
        width = 13, height = 15)


ggsave("output/Fig02_Core&CRAT_wNUM.png",
       p,
       width = 13, height = 15)

ggsave("output/Fig02_Core&CRAT_wNUM.pdf",
       p,
       width = 13, height = 15,
       useDingbats=FALSE, limitsize=FALSE)

```


